A Systematic Literature Review on Extensions to the Role-Based Access Control Reference Model

### 2014-02-15 paper/sections/categorization.tex
Updates.

During the paper reading phase, we identify classification themes, which are central topic or context of the RBAC extension models. 
Our classification process yields a set of categories based upon observations during the paper reading phase.
For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
We classify this paper into the Privacy category. 
Some papers presented this direct pronouncement of theme of the RBAC extension models, whereas others were less obvious.
Thus, we developed a set of guidelines to aide in determining a set of eight categories. 
In developing these guidelines, we defined each category by a single noun-phrase descriptor.
The guidelines were as follows:

@@

The previous example paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name. 
However, in reading the body of the paper, we determined that this paper discussed RBAC extension based on context. 
In Section 4.1, we offer definitions for each of the eight observed categories.

@@

We provide a definition for each observed model category based on data extracted from the primary sources and the English definitions for each noun-phrase.
Some category descriptors contain abbreviations in parenthesis that match the shortened name found in Table 6 (within Appendix 1) \ref{tab:categorization} which presents each primary source within it's designated category.

@@

  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies. For example, specific locations permit roles to conduct actions from 8:00 a.m. to 5:00 p.m.

@@

\greybox{We found eight categories that exist within the RBAC extension models: Constraint, Context, 
Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other
seven categories fall under the category of Context.}

>>>>

During the paper reading phase, we identified that a variety of common terms were emerging to describe the RBAC extension models. 
We developed a process by which to build a classification of the RBAC extension models based upon observations during the paper reading phase.
For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
Some papers appeared to present a direct pronouncement of their classification, whereas others were less obvious.
Thus, we developed a set of guidelines to aide in determining a set of eight categories. 
In developing these guidelines, we defined each category by a single noun-phrase descriptor.
The guidelines were as follows:

@@

The previous example paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name. 
However, in reading the body of the paper, we determined that the paper discussed RBAC extension based on context. 
In Section 4.1, we offer definitions for each of the eight observed categories.

@@

We provide a definition for each observed model category based on data extracted from the primary sources and the English definitions for each noun-phrase.
Some category descriptors contain abbreviations in parenthesis that match the shortened name found in Table 6 (within Appendix 1) which presents each primary source within it's designated category.

@@

  \item \textbf{Spatio-Temporal}: The extension model combines the use of either or both spatial (location-based) and temporal (time-based) constraints in specifying access control policies. For example, specific locations permit roles to conduct actions from 8:00 a.m. to 5:00 p.m.

@@

\greybox{We found eight categories that exist within the RBAC extension models: Constraint, Context, Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. The context category was a superset of all other categories since all categories were found to consider context.}

### 2014-02-15 paper/sections/core_rbac.tex
Updates.

RBAC provides effective and efficient permissions management for operations, especially when sharing resources within an organization.
Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
of RBAC existed among practitioners or within the research community. 
Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
This lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
Moreover, in cases where organizations were concerned with adopting RBAC,
evaluation and comparison of RBAC technologies developed by different vendors was difficult.
NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for 
RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.

NIST's work can directly benefit organizations by lowering the cost of early phase research and development, and implementation of RBAC.
Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard saved \$1.1 billion.

@@

In RBAC, a user can exercise a permission only if the user is assigned to a role that contains the permission.
In addition to the four basic entities, two functions are defined:
user assignment ($UA$) and permission assignment ($PA$) functions.
$UA$ represents assignment of users to roles.
$PA$ represents assignment of permissions to roles.

\begin{figure}[ht]
    \centering
        \includegraphics[width=4.0in]{sections/core-model.png}
    \caption{\label{fig:overview}Diagram in Core RBAC reference model\cite{ferraiolokuhn}.}
\end{figure}

Figure~\ref{fig:overview} presents an overview of the core RBAC reference model where entities and their relations are described.
Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
Permissions are associated with possible users' pre-defined operation on an object (e.g., execute a file).
Note that, at user or role activation, a session associated with user or role is established.

@@

The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. 
Separation of duty relations enforce conflicts of interest among roles. This model defines two types of separation of duty relations: static and dynamic.

>>>>

RBAC provides effective and efficient permissions management for operations, especially when sharing resources within an organization.
Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
of RBAC existed among practitioners or within the research community. 
Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
The lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
Moreover, in cases where organizations were concerned with adopting RBAC,
evaluation and comparison of RBAC technologies developed by different vendors was difficult.
NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for 
RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.

NIST's work can directly benefit organizations by lowering the cost of early phase research and development, and implementation of RBAC.
Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
The analysts estimate that the use of RBAC technology has saved industry \$6.1 billion from 1992 to 2009 and the research and development work by NIST to produce an RBAC standard has saved an additional \$1.1 billion that wold have otherwise been spent across the industry doing research and development on RBAC.

@@

In RBAC, a user can exercise a permission only if the user is assigned to a role that contains the permission.
In addition to the four basic entities, two functions are defined:
user assignment ($UA$) and permission assignment ($PA$) functions.
$UA$ represents assignment of users to roles.
$PA$ represents assignment of permissions to roles.
Permissions are associated with possible users' pre-defined operation on an object (e.g., execute a file).
Note that, at user or role activation, a session associated with user or role is established.

@@

The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. 
Separation of duty relations enforce conflicts of interest among roles. 
The Constrained RBAC model defines two types of separation of duty relations: static and dynamic.

### 2014-02-15 paper/sections/process.tex
Updates.

We adopted and applied a systematic literature review process following recommendations from Kitchenham's suggested processes~\cite{kitchenham2007guidelines}.  The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage.  The stages were as follows:

@@

For the first phase of our systematic literature review, we developed a search strategy for finding papers.  The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search terms.  The search was performed by applying each search term to each engine incrementally until the stopping criteria were met. Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination.  The search algorithm was performed as follows:

@@

The result set size stopping criteria was chosen due to a technical limitation of some search engines.  The stopping criteria related to the last ten titles are meant to stop after relevant results are no longer being returned by the search engine.  After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
Out of the master list of 1,716 papers, two reviewers, denoted by Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources. Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.

\begin{table}
\centering
\caption{Paper counts after applying search strategy}
\vspace{0.1 in}
\begin{tabular}{ | l | r | r | r | r | }

@@

The title elimination round resulted in Reviewer 1 selecting 305 papers, and Reviewer 2 selecting 176 papers with 149 papers of overlap between the two. 
There were 332 papers found not to be in common, of which, 149 were rejected and 141 retained after a second review.

@@

The second round of elimination was based on reading of the abstracts of papers that survived title elimination.  Researchers read each abstract and evaluated relevancy based off:

@@

The abstract elimination round resulted in Reviewer 1 selecting 86 papers, and Reviewer 2 selecting 102 papers with 51 papers of overlap between the two. 
There were 137 papers found not to be in common, of which, 116 were rejected and 21 retained after a second review.

@@

The content elimination round resulted in Reviewer 1 selecting 46 papers, and Reviewer 2 selecting 42 papers with 24 papers of overlap between the two. 
Between the two reviewers selections, there were 64 papers not in common, of which, 59 were rejected and 5 retained after a second review.

@@

After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends. 
The first step was to take the individual data generated from the final elimination round and organize this information around the research questions. 
During the paper reading round and resulting data, the fact that the papers were falling into a number of categorizations became evident. 
Thus, the first step undertaken was to answer the question of what categories exist for the RBAC extension models and what papers fell into what categories.

>>>>

We adopted and applied a systematic literature review process following recommendations from Kitchenham and Charter's suggested processes~\cite{kitchenham2007guidelines}. 
The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage. 
The stages were as follows:

@@

For the first phase of our systematic literature review, we developed a search strategy for finding papers. 
The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search terms. 
The search was performed by applying each search term to each engine incrementally until the stopping criteria were met. 
Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination. 
The total column represents the total for each individual search engine with a grand total across all search engines. 
The combined total represents the net paper total after removing duplicate entries. 
The search algorithm was performed as follows:

@@

The result set size stopping criteria was chosen due to a technical limitation of some search engines. 
The stopping criteria related to the last ten titles are meant to stop after relevant results are no longer being returned by the search engine. 
After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
Out of the master list of 1,716 papers, two reviewers conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
The two reviewers are denoted by Author 1 and Author 2 where the former is the first author on the paper (Eric) and the latter is the second author on the paper (JeeHyun).
Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.

\begin{table}
\centering
\caption{Paper counts after applying Step 1}
\vspace{0.1 in}
\begin{tabular}{ | l | r | r | r | r | }

@@

The title elimination round resulted in Author 1 selecting 305 papers, and Author 2 selecting 176 papers with 149 papers of overlap between the two. 
All 149 papers found to be in both reviewers lists were kept. 
The 332 papers found not to be in common were then slated for a second round of review. 
The second round of review consisted of the reviewers discussing their perception of each paper title as the title related to the elimination criteria and making a joint decision to keep or reject. 
The second review resulted in 149 rejections and 141 being retained.

@@

The second round of elimination was based on reading of the abstracts of papers that survived title elimination. 
The reviewers read each abstract and evaluated relevancy based off:

@@

The abstract elimination round resulted in Author 1 selecting 86 papers, and Author 2 selecting 102 papers with 51 papers of overlap between the two. 
All 51 papers found to be in both reviewers lists were kept. 
The 137 papers found not to be in common were then slated for a second round of review. 
The second round of review consisted of the reviewers discussing their perception of each the paper abstract as each related to the elimination criteria and making a joint decision to keep or reject based on the joint outcome of the discussion.
The second review resulted in 116 rejections and 21 being retained.

@@

The content elimination round resulted in Author 1 selecting 46 papers, and Author 2 selecting 42 papers with 24 papers of overlap between the two. 
Between the two reviewers selections, there were 64 papers not in common, of which, 59 were rejected and 5 retained after a second review.

@@

After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions to look for trends. 
The first step was to take the individual data generated from the final elimination round and organize this information around the research questions. 
During the paper reading round and resulting data, the fact that the papers were falling into a number of categorizations became evident. 
Thus, the first step undertaken was to answer the question of what categories exist for the RBAC extension models and what papers fell into what categories.
